---
title: "PhD thesis - analyses and figures"
author: "Pierre Priou - ppr@akavaplan.niva.no"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: true
---

This document contains the code used to create figures for my PhD thesis. 

# Packages required

Load packages.

```{r packages, message=FALSE}
library(raster)
library(sf)
library(tidyverse)
library(lubridate)
library(marmap)
library(cmocean)
library(ggnewscale)
source("R/getNOAA.ice.bathy.R")
```

Set figure theme.

```{r update-fig-theme}
# Theme for figures
theme_set(theme_bw())
theme_update(panel.grid = element_blank(), 
             panel.border = element_blank(), 
             axis.line = element_line(),
             axis.text = element_text(size = 8),
             axis.title = element_text(size = 9),
             legend.title = element_text(size = 9),
             strip.text.x = element_text(size = 8, face = "plain", hjust = 0.5),
             strip.background = element_rect(colour = "transparent",  fill = "transparent"),
             legend.margin = margin(0, 0, 0, 0),
             legend.box.margin = margin(0, 0, -8, 0),
             plot.margin = unit(c(0.05, 0.05, 0.05, 0.05), "in"))
```

# Introduction

## The Arctic Ocean

### Fig 1.1. Map Arctic Ocean

Map of the Arctic Ocean with bathymetry. 

Define new projection.

```{r}
# Polar stereographic projection for bathy and all other files
arctic_6931 <- raster(extent(-3500, 3500, -3500, 3500), crs = "EPSG:6931") 
projection(arctic_6931) <- gsub("units=m", "units=km",  projection(arctic_6931))
```

Load data: coastline, rivers, graticules, and bathymetry.

```{r ,warning=FALSE}
# Coastline shapefiles
coast_6931 <- rgdal::readOGR("data/bathy/ne_10m_land.shp", verbose = F) %>% 
  spTransform(CRSobj = crs("EPSG:4326")) %>% 
  crop(extent(-180, 180, 0, 90)) %>% # Crop shapefile
  spTransform(CRSobj = crs(arctic_6931)) %>% # Project shapefile in laea
  fortify() %>% # Convert to a dataframe for ggplot
  rename(xc = long, yc = lat)

# Rivers
rivers_6931 <- rgdal::readOGR("data/bathy/ne_10m_rivers_lake_centerlines.shp", verbose = F) %>% 
  spTransform(CRSobj = crs("EPSG:4326")) %>% 
  crop(extent(-180, 180, 0, 90)) %>% # Crop shapefile
  spTransform(CRSobj = crs(arctic_6931)) %>% # Project shapefile in laea
  fortify() %>% # Convert to a dataframe for ggplot
  rename(xc = long, yc = lat)

# Longitudes
lon_6931 <- data.frame(lon = rep(seq(-180, 180, 45), each = 200)) %>%
  group_by(lon) %>%
  mutate(lat =  seq(0, 90, length.out = 200)) %>%
  ungroup() %>%
  # Transform into sf
  st_as_sf(coords = c("lon", "lat"), crs = st_crs("EPSG:4326"), remove = F) %>% 
  # Change projection
  st_transform(crs = st_crs(arctic_6931)) %>%
  mutate(xc = st_coordinates(.)[,1],
         yc = st_coordinates(.)[,2], 
         lon = factor(lon)) %>%
  st_drop_geometry()

# Latitudes
lat_6931 <- data.frame(lat = rep(seq(0, 90, 10), each = 200)) %>%
  group_by(lat) %>%
  mutate(lon =  seq(-180, 180, length.out = 200)) %>%
  ungroup() %>%
  # Transform into sf
  st_as_sf(coords = c("lon", "lat"), crs = st_crs("EPSG:4326"), remove = F) %>% 
  # Change projection
  st_transform(crs = st_crs(arctic_6931)) %>%
  mutate(xc = st_coordinates(.)[,1],
         yc = st_coordinates(.)[,2], 
         lon = factor(lon)) %>%
  st_drop_geometry()

# Arctic circle 
arctic_circle_6931 <- data.frame(lat = rep(66.5, each = 200)) %>%
  group_by(lat) %>%
  mutate(lon =  seq(-180, 180, length.out = 200)) %>%
  ungroup() %>%
  # Transform into sf
  st_as_sf(coords = c("lon", "lat"), crs = st_crs("EPSG:4326"), remove = F) %>% 
  # Change projection
  st_transform(crs = st_crs(arctic_6931)) %>%
  mutate(xc = st_coordinates(.)[,1],
         yc = st_coordinates(.)[,2], 
         lon = factor(lon)) %>%
  st_drop_geometry()

# Bathymetry
bathy_6931 <- read_csv("data/bathy/bathy_CAO_proj6931.csv", show_col_types = F)
```

Map of the CAO and its surrounding seas.

```{r}
map_AO <- bathy_6931 %>%
  mutate(depth = if_else(depth <= 0, depth, NaN)) %>% 
  ggplot(aes(x = xc, y = yc)) +
  # Plot bathy
  geom_raster(aes(fill = depth), alpha = 0.6) +
  scale_fill_cmocean("Depth (m)",  name = "deep", direction = -1, limits = c(-4500, 0), oob = scales::squish,
                     # breaks = c(-4000, -3000, -2000, -1000, -500, -200, 0),
                     # labels = c(4000, 3000, 2000, 1000, 500, 200, 0)) +
                     breaks = seq(-4000, 0, 500), labels = seq(-4000, 0, 500)*-1,
                     alpha = 0.6) +
  # scale_fill_grey("Depth (m)", start = 0.9, end = 0.3) + 
  # guides(fill = "none") + 
  new_scale_fill() +
  # Plot 200 m isobath
  # geom_contour(aes(z = depth), breaks = -200, col = "black", lwd = 0.1, alpha = 0.8) +
  # Plot coastlines
  # geom_polygon(data = coast_6931, aes(x = xc, y = yc, group = group), fill = "grey30") +
  geom_polygon(data = coast_6931, aes(x = xc, y = yc, group = group), fill = "grey40", lwd = 0) +
  # Graticules 
  geom_path(data = lat_6931, aes(x = xc, y = yc, group = lat), col = "grey80", lwd = 0.1, alpha = 0.5) +
  geom_path(data = lon_6931, aes(x = xc, y = yc, group = lon), col = "grey80", lwd = 0.1, alpha = 0.5) +
  # Arctic circle
  # geom_path(data = arctic_circle_6931, aes(x = xc, y = yc), col = "white", lty = 2, linewidth = 0.4) +
  coord_fixed(xlim = c(-4500, 4500), ylim = c(-4500, 4500), expand = F) +
  guides(fill = "none") +
  # Plot North Pole
  # geom_point(aes(x = 0, y = 0), size = 0.05) +
  theme(legend.position = "right", 
        legend.key.height = unit(0.4, "in"),
        legend.key.width = unit(0.1, "in"),
        panel.border = element_rect(fill = NA),
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())
ggsave("plots/base_map_AO.png", map_AO, height = 4, width = 6.5, dpi = 600, units = "in")

# Zoom on CAO
map_CAO <- bathy_6931 %>%
  mutate(depth = if_else(depth <= 0, depth, NaN)) %>% 
  ggplot(aes(x = xc, y = yc)) +
  # Plot bathy
  geom_raster(aes(fill = depth), alpha = 0.8) +
  scale_fill_cmocean("Depth (m)",  name = "deep", direction = -1, limits = c(-4500, 0), oob = scales::squish,
                     # breaks = c(-4000, -3000, -2000, -1000, -500, -200, 0),
                     # labels = c(4000, 3000, 2000, 1000, 500, 200, 0)) +
                     breaks = seq(-4000, 0, 500), labels = seq(-4000, 0, 500)*-1) +
  # scale_fill_grey("Depth (m)", start = 0.9, end = 0.3) + 
  # guides(fill = "none") + 
  new_scale_fill() +
  # Plot 200 m isobath
  # geom_contour(aes(z = depth), breaks = -200, col = "black", lwd = 0.1, alpha = 0.8) +
  # Plot coastlines
  # geom_polygon(data = coast_6931, aes(x = xc, y = yc, group = group), fill = "grey30") +
  geom_polygon(data = coast_6931, aes(x = xc, y = yc, group = group), fill = "grey40") +
  # geom_path(data = rivers_6931, aes(x = xc, y = yc, group = group), col = "grey65", lwd = 0.2) +
  # Graticules 
  geom_path(data = lat_6931, aes(x = xc, y = yc, group = lat), col = "grey80", lwd = 0.1, alpha = 0.5) +
  geom_path(data = lon_6931, aes(x = xc, y = yc, group = lon), col = "grey80", lwd = 0.1, alpha = 0.5) +
  # Arctic circle
  geom_path(data = arctic_circle_6931, aes(x = xc, y = yc), col = "white", lty = 2, linewidth = 0.4) +
  coord_fixed(xlim = c(-3100, 2700), ylim = c(-3000, 3200), expand = F) +
  guides(fill = "none") +
  # Plot North Pole
  # geom_point(aes(x = 0, y = 0), size = 0.05) +
  theme(legend.position = "right", 
        legend.key.height = unit(0.4, "in"),
        legend.key.width = unit(0.1, "in"),
        panel.border = element_rect(fill = NA),
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())
ggsave("plots/base_map_CAO.png", map_CAO, height = 4, width = 6.5, dpi = 600, units = "in")
# map_CAO
```

Other map AO, black with no background.

```{r}
map_AO_black <- coast_6931 %>%
  ggplot() +
  # Plot coastlines
  geom_polygon(aes(x = xc, y = yc, group = group), fill = "grey40") +
  coord_fixed(xlim = c(-4500, 4500), ylim = c(-4500, 4500), expand = F) +
  guides(fill = "none") +
  # Plot North Pole
  # geom_point(aes(x = 0, y = 0), size = 0.05) +
  theme(legend.position = "right", 
        legend.key.height = unit(0.4, "in"),
        legend.key.width = unit(0.1, "in"),
        panel.border = element_rect(fill = NA),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA), 
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())
ggsave("plots/base_map_AO_black.png", bg = "transparent", map_AO_black, height = 4, width = 6.5, dpi = 600, units = "in")
map_AO_black
```

## Light comfort zone

Load extrapolated irradiance data from TriOS. 

```{r light-extrap-loading}
light <- read_csv("data/light/Ed_integrated_extrapolated.csv", show_col_types = F, locale = locale(tz = "UTC")) %>%
  # Use date for combining with other dataset
  rename(date_loc = date_round) %>%
  # Convert from local time (Quebec) to UTC
  mutate(date = date_loc + hours(4),
         day = day(date_loc)) %>%
  mutate(alpha = if_else(between(ed_log_PAR, -9, -6), "in", "out")) %>%
  filter(is.na(station_group) == F) 

LCZ <- light %>%
  filter(day > 18) %>%
  ggplot() +
  geom_tile(aes(x = date_loc, y = depth, alpha = alpha), fill = "grey", na.rm = T) + 
  geom_contour(aes(x = date_loc, y = depth, z = ed_log_PAR), breaks = c(-6, -9), col = "red", na.rm = T, lwd = 0.4) +
  scale_alpha_discrete(range = c(1, 0)) +
  scale_y_reverse() + 
  theme(legend.position = "right", 
        legend.key.height = unit(0.4, "in"),
        legend.key.width = unit(0.1, "in"),
        panel.border = element_rect(fill = NA),
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA), 
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())
ggsave("plots/LCZ.png", bg = "transparent", LCZ, height = 4, width = 6.5, dpi = 600, units = "in")
```




